---
title: "Recreating UKBB FI in R"
author: "Ruth CE Bowyer"
date: "2023-08-30"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
```

## 0. About

Converting the Stata script provided by Williams et al., to generate an FI in UKBB

Differences are highlighted below but in brief:


This script was run mostly via CREATE

```{r}

library(dplyr)

```

## 1. Data

Variables to select as per William's et al.,:

Variables required from UKB data (these represent data field numbers, along with the Stata prefix for each variable 
	('s_' or 'n_' for string or numeric vars, respectively):
n_eid  		-- participant identifier x
n_31_0* 	-- Sex x
n_34_* 		-- Year of birth x
n_52_* 		-- Month of birth x
n_21003_0*  -- Age when attended assessment centre x
n_21000* 	-- Ethnicity x
n_1200_0* 	-- Sleeplessness / insomnia x
n_6148_0* 	-- Eye problems / disorders x
n_2227_0*  	-- Other eye problems x
n_2020_0* 	-- Loneliness / isolation x
n_2247_0* 	-- Hearing difficulties x
n_2188_0* 	-- Long-standing illness, disability or infirmity x
n_2178_0* 	-- Overall self-rated health x
n_2296_0* 	-- Falls in last year x
n_2316_0* 	-- Chest wheeze or whistling x
n_2080_0* 	-- Tiredness / Lethargy x
n_6150_0* 	-- Vascular problems	x
n_20002_0* 	-- Non-cancer illness codes (all array variables are necessary to keep) x
n_6149_0* 	-- Mouth / dental problems x 
n_2335_0* 	-- Chest pain x
n_2443_0* 	-- Diabetes x
n_2453_0* 	-- Cancer diagnosis x
n_134_0* 	-- Number of self-reported cancers x
n_2463_0* 	-- fractured/broken bones x
n_6152_0*	-- Multiple medical conditions reported - e.g. DVT, asthma, eczema x
n_3786_0*	-- Age of asthma diagnosis x
n_2473_0* 	-- Other serious medical conditions reported x
n_2050_0*  	-- Depressed moods x
n_1970_0* 	-- Nervous feelings x
n_1930_0* 	-- Miserableness x
n_6159_0* 	-- Pain in last month x
n_136_0* 	-- Number of operations x

n_6177_0* 	-- Medications used for cholesterol, blood pressure or diabetes (male)
n_6153_0* 	-- Medications used for cholesterol, blood pressure or diabetes (female)
n_137_0* 	-- Number of medications used

n_40007_* 	-- Age at death for deceased participants

```{r eval=FALSE, include=T}
# This chunk ran for set up and then just loaded as below for subsequent iterations
#df.a derived as follows: 

load("~/max2ruth2/bd671644_recoded.Robj")

df.a <- bd %>% select("f.eid", "f.31.0.0", "f.34.0.0", "f.52.0.0",
                      "f.21003.0.0":"f.21003.3.0", "f.21000.0.0":"f.21000.3.0",
                      "f.1200.0.0":"f.1200.3.0", "f.6148.0.0":"f.6148.3.4", 
                      "f.2227.0.0":"f.2227.3.0", "f.2020.0.0":"f.2020.3.0",
                      "f.2247.0.0":"f.2247.3.0",
                      "f.2188.0.0":"f.2188.3.0", "f.2178.0.0":"f.2178.3.0",
                      "f.2296.0.0":"f.2296.3.0", "f.2316.0.0":"f.2316.3.0",
                      "f.2080.0.0":"f.2080.3.0", "f.6150.0.0":"f.6150.2.3",
                      "f.20002.0.0":"f.20002.3.33", "f.6149.0.0":"f.6149.3.5",
                      "f.2335.0.0":"f.2335.3.0", "f.2443.0.0":"f.2443.3.0",
                      "f.2453.0.0":"f.2453.3.0", "f.134.0.0":"f.134.3.0",
                      "f.2463.0.0":"f.2463.3.0", "f.6145.0.0":"f.6145.3.2",
                      "f.3786.0.0":"f.3786.3.0", "f.2473.0.0":"f.2473.3.0", "f.2050.0.0":"f.2050.3.0",
                      "f.1970.0.0":"f.1970.3.0", "f.1930.0.0":"f.1930.3.0", "f.6159.0.0":"f.6159.3.6",
                      "f.136.0.0":"f.136.3.0", "f.6177.0.0":"f.6177.3.2", "f.6153.0.0":"f.6153.3.3",
                      "f.137.0.0":"f.137.3.0", "f.40007.0.0":"f.40007.1.0",
                      "f.6152.0.0":"f.6152.3.4")

write.csv(df.a,"/users/k1511199/UKBB_multimorbidity/UKBB.df.a.csv", row.names=F)

```

Many of the array vals seem to be completely missing NA 

```{r}

df.a <- read.csv("/users/k1511199/UKBB_multimorbidity/UKBB.df.a.csv")
## NA those who withdrew 
x<-read.csv("~/max2ruth/w90865_2023-04-25.csv",h=F)

df.a <- df.a[-which(df.a$f.eid%in%x$V1),]

## set row names as study id for sense checking

row.names(df.a) <- paste0("id",df.a$f.eid)
```


## 2. Write functions 

Recoding NA to -99999 as, presumably due to package version install on CREATE and dplyr, NA and ifelse behaviour is unusual


```{r recode function}

#Function to search within a range of cols for a text variable and recode as 1 - returns a vector of the var 

text_recode <- function(var1, #first var in range
                        var2 = NULL, #second var in range
                        text, #Text to search in array and recode to 1
                        NAvals, ...){ #Text to recode as NA (-99999) sep multiple with |
  var1 = var1
  var2 = var2
  text = text
  df.b <- df.a %>% select(var1:var2)
  
  df.b <- apply(df.b,2,as.character) #Can throw error in next step if all factors instead
  
  df.b[is.na(df.b)] <- -99999
  
  var <- apply(df.b,1, function(b){
    v <- ifelse(any(grepl(text, b)), 1, 
           ifelse(all(grepl(NAvals, b)|grepl(-99999, b)), -99999, 0))})
  return(var)}

#Function as above but where the question only allows inference of not having the condition - so all no or NA answers are recorded as 0
text_recode0only <- function(var1, #first var in range
                        var2 = NULL, #second var in range
                        text, #Text to search in array and recode to 0
                        NAvals, ...){ #Text to recode as NA (-99999), sep multiple with |
  var1 = var1
  var2 = var2
  text = text
  df.b <- df.a %>% select(var1:var2)
  
  df.b[is.na(df.b)] <- -99999
  
    var <- apply(df.b,1, function(b){
    v <- ifelse(any(grepl(text, b)), 0, -99999)})
  return(var)}


#Function for numeric variables -- codes any mention of morb as 1, and if no 1s and all responses arent NA, returns 0

num_recode <- function(var1, #first var in range
                        var2 = NULL, #second var in range
                        code){ #Code for morb 
  var1 = var1
  var2 = var2
  code = code
  df.b <- df.a %>% select(var1:var2)
  

  df.b[is.na(df.b)] <- -99999
  var <- apply(as.data.frame(df.b),1,function(b){
    b <- as.numeric(b)
    ifelse(all(b==-99999), -99999, 
               ifelse(any(b==code),1,0))
  })
  
  return(var)}


```

## 3. Clean datasets 


### 3.1 Glaucoma

- Within f.6148* - text var
- is code 1277 in 20002 
- Recode do not know and pref not to answer to NA 
- Var 2227 - other eye problems - Williams included 'No' as 0 so will do that here too

Questions from og methodology: 

I will recode pref not to answer and don't know to NA as in the og -- but im not sure this is correct given the MICE? (and some don't knows could be recoded to 0)

'	* recoding 6148 'do not know' respondents to 0 if they were interviewed and still had no record of glaucoma:
replace i1_glaucoma = 0 if i1_!=1 & n_20002_0_0!=. & n_6148_0_0==-1 '

I'm not sure what variable is meant by this or the variable i1 - which i think is to do with interviewing but I can't see a corresponding var in the html, or which variable this relates to 

99999 in 20002 -- unclassifable so I don't think coding as NA is appropiate but discuss
https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=6

Where all responses in 20002 is NA - the fnc codes as NA(but could be 0 if the NAs are responses vs not being asked)


```{r message = F, warning=F}

glaucoma_t <- text_recode(var1= "f.6148.0.0", #first var in range
                        var2 = "f.6148.3.4",#second var in range
                        text = "Glaucoma", #Text to search in array and recode to 1
                        NAvals = "Prefer not to answer|Do not know") #Text to recode, sep multiple with 

table(glaucoma_t, useNA="ifany")

glaucoma_n <- num_recode(var1= "f.20002.0.0", #first var in range
                        var2 = "f.20002.3.33",#second var in range
                        code=1277) #Text to search in array and recode to NA)

table(glaucoma_n, useNA="ifany")

#This Q for no other eye problems
noeyeprobs <- text_recode0only(var1="f.2227.0.0",
                               var2="f.2227.3.0",
                               text="No")


glaucoma <- ifelse(glaucoma_n==1|glaucoma_t==1,1, #This condition first
                ifelse(glaucoma_n==0|glaucoma_t==0|noeyeprobs==0,0,NA))

table(glaucoma, useNA="ifany")
```

### 3.2 Cataracts

Deviation from DW method - for this var if answer is Do not know, coding as 0 (as if they don't know they probably don't have cataracts)

```{r message = F, warning=F}

cataracts_t <- text_recode(var1= "f.6148.0.0", #first var in range
                        var2 = "f.6148.3.4",#second var in range
                        text = "Cataract", #Text to search in array and recode to 1
                        NAvals = "Prefer not to answer") #Text to recode, sep multiple with 

table(cataracts_t, useNA="ifany")

cataracts_n <- num_recode(var1= "f.20002.0.0", #first var in range
                        var2 = "f.20002.3.33",#second var in range
                        code=1278) #Text to search in array and recode to NA)

table(cataracts_n, useNA="ifany")


#same no other eye probs logic

cataracts <- ifelse(cataracts_n==1|cataracts_t==1,1, #This condition first
                ifelse(cataracts_n==0|cataracts_t==0|noeyeprobs==0,0,NA))

table(cataracts, useNA="ifany")
```

### 3.3 Hearing
 - Hearing: dichotomous on any/no hearing problems; deaf also coded to 1

Those answering 'prefer not to answer'  are coded as missing 
- Deviating from DW if responded 'Do not know' coded as 0

```{r}
hearing_t <-  text_recode(var1= "f.2247.0.0", #first var in range
                        var2 = "f.2247.3.0",#second var in range
                        text = "Yes", 
                        NAvals = "Prefer not to answer") #Text to recode, sep multiple with 

table(hearing_t, useNA="ifany")

hearing <- ifelse(hearing_t==-99999, NA, hearing_t)
table(hearing, useNA="ifany")
```
### 3.4 Infirmity 
	
Coding as in 3c
	
```{r}
infirmity_t <-  text_recode(var1= "f.2188.0.0", #first var in range
                        var2 = "f.2188.3.0",#second var in range
                        text = "Yes", 
                        NAvals = "Prefer not to answer") #Text to recode, sep multiple with 

table(infirmity_t, useNA="ifany")

infirmity <- ifelse(infirmity_t==-99999, NA, infirmity_t)
table(infirmity, useNA="ifany")

```

### 3.5 self-rated health

(DW's STATA code:)
gen i5_self_health = 0 if n_2178==1
replace i5 = 0.25 if n_2178==2
replace i5 = 0.5 if n_2178==3
replace i5 = 1 if n_2178==4 

**Q:**Something to consider is timing - I don't know enough about these vars (ie are they from consecutive Qs and which case shouldnt for something like this we take the most recent? as it can fluctuate?) For now, its coded as the worst reported 
	

```{r}
  var1 = "f.2178.0.0"
  var2 = "f.2178.3.0"
  text0 = "Excellent"
  text0.25 = "Good"
  text0.5 = "Fair"
  text1= "Poor"
  NAvals = "Do not know|Prefer not to answer"
  
  df.b <- df.a %>% select(var1:var2)
  
  df.b[is.na(df.b)] <- -99999
  
SRH_t <- apply(df.b,1, function(b){
    v <- ifelse(any(grepl(text1, b)), 1,
                ifelse(any(grepl(text0.5, b)), 0.5,
                       ifelse(any(grepl(text0.25, b)), 0.25,
                              ifelse(any(grepl(text0, b)), 0, -99999))))})

table(SRH_t, useNA="ifany")

SRH <- ifelse(SRH_t==-99999, NA, SRH_t)
table(SRH, useNA="ifany")

```
### 3.6 Falls

(DW's STATA code):

gen i6_falls = 0 if n_2296==1
replace i6_falls = 0.5 if n_2296==2
replace i6_falls = 1 if n_2296==3



```{r}

  var1 = "f.2296.0.0"
  var2 = "f.2296.3.0"
  text0 = "No falls"
  text0.5 = "Only one fall"
  text1= "More than one fall"
  NAvals = "Prefer not to answer"
  
  df.b <- df.a %>% select(var1:var2)
  
  df.b[is.na(df.b)] <- -99999
  
falls_t <- apply(df.b,1, function(b){
    v <- ifelse(any(grepl(text1, b)), 1,
                ifelse(any(grepl(text0.5, b)), 0.5,
                              ifelse(any(grepl(text0, b)), 0, -99999)))})

table(falls_t, useNA="ifany")

falls <- ifelse(falls_t==-99999, NA, falls_t)
table(falls, useNA="ifany")


```

### 3.7 Breathing/wheeze

gen i7_wheeze = 1 if n_2316==1
replace i7 = 0 if n_2316==0


```{r}

wheeze_t <- text_recode(var1= "f.2316.0.0", #first var in range
                        var2 = "f.2316.3.0",#second var in range
                        text = "Yes", #Text to search in array and recode to 1
                        NAvals = "Prefer not to answer|Do not know") 

table(wheeze_t, useNA="ifany")

wheeze <- ifelse(wheeze_t==-99999, NA, wheeze_t)
table(wheeze, useNA="ifany")


```
	
### 3.8 Fatigue

gen i8_fatigue = 0 if n_2080==1 
replace i8 = 0.25 if n_2080==2
replace i8 = 0.75 if n_2080==3
replace i8 = 1 if n_2080==4


```{r}

  var1 = "f.2080.0.0"
  var2 = "f.2080.3.0"
  text0 = "Not at all"
  text0.25 = "Several days"
  text0.75 = "More than half the days"
  text1= "Nearly every day"
  NAvals = "Do not know|Prefer not to answer"
  
  df.b <- df.a %>% select(var1:var2)
  
  df.b[is.na(df.b)] <- -99999
  
fatigue_t <- apply(df.b,1, function(b){
    v <- ifelse(any(grepl(text1, b)), 1,
                ifelse(any(grepl(text0.75, b)), 0.75,
                       ifelse(any(grepl(text0.25, b)), 0.25,
                              ifelse(any(grepl(text0, b)), 0, -99999))))})

table(fatigue_t, useNA="ifany")

fatigue <- ifelse(fatigue_t==-99999, NA, fatigue_t)
table(fatigue, useNA="ifany")

```



### 3.9 Myocardial infarction


		* recoding heart attack reported as '1075' in field 20002
		
forvalues n =  0(1)28 {
replace i9 = 1 if n_20002_0_`n'==1075
}
		* NB: no one coded as 'do not know' from field 6150 to recode here

```{r}
MI_t <- text_recode(var1= "f.6150.0.0", #first var in range
                        var2 = "f.6150.2.3",#second var in range
                        text = "Heart attack", #Text to search in array and recode to 1
                        NAvals = "Prefer not to answer") #Text to recode, sep multiple with 

table(MI_t, useNA="ifany")

MI_n <- num_recode(var1= "f.20002.0.0", #first var in range
                        var2 = "f.20002.3.33",#second var in range
                        code=1075) #Text to search in array and recode to NA)
table(MI_n, useNA="ifany")


MI <- ifelse(MI_n==1|MI_t==1,1, #This condition first
                ifelse(MI_n==0|MI_t==0,0,NA))

table(MI, useNA="ifany")


```

### 3.10 Angina 

```{r}
angina_t <- text_recode(var1= "f.6150.0.0", #first var in range
                        var2 = "f.6150.2.3",#second var in range
                        text = "Angina", #Text to search in array and recode to 1
                        NAvals = "Prefer not to answer") #Text to recode, sep multiple with 

table(angina_t, useNA="ifany")

angina_n <- num_recode(var1= "f.20002.0.0", #first var in range
                        var2 = "f.20002.3.33",#second var in range
                        code=1074) #Text to search in array and recode to NA)
table(angina_n, useNA="ifany")


angina <- ifelse(angina_n==1|angina_t==1,1, #This condition first
                ifelse(angina_n==0|angina_t==0,0,NA))

table(angina, useNA="ifany")


```

### 3.11 Stroke

	* recoding stroke reported as '1081' (ambiguous stroke) or '1583' (ischaemic) in field 20002

forvalues n =  0(1)28 {
replace i11 = 1 if n_20002_0_`n'==1081 | n_20002_0_`n'==1583
}


```{r}

stroke_t <- text_recode(var1= "f.6150.0.0", var2 = "f.6150.2.3",
                        text = "Stroke", NAvals = "Prefer not to answer") 

table(stroke_t, useNA="ifany")

#'1081' (ambiguous stroke)
stroke_n1 <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1081)
table(stroke_n1, useNA="ifany")

#'1583' (ischaemic)
stroke_n2 <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1583) #Text to search in array and recode to NA)
table(stroke_n2, useNA="ifany")

stroke <- ifelse(stroke_n1==1|stroke_n2==1|stroke_t==1,1, #This condition first
                ifelse(stroke_n1==0|stroke_n2==0|stroke_t==0,0,NA))

table(stroke, useNA="ifany")



```

### 3.12 High blood pressure (HBP)


```{r}
HBP_t <- text_recode(var1= "f.6150.0.0", #first var in range
                        var2 = "f.6150.2.3",#second var in range
                        text = "High blood pressure", #Text to search in array and recode to 1
                        NAvals = "Prefer not to answer") #Text to recode, sep multiple with 

table(HBP_t, useNA="ifany")

HBP_n <- num_recode(var1= "f.20002.0.0", #first var in range
                        var2 = "f.20002.3.33",#second var in range
                        code=1065) #Text to search in array and recode to NA)
table(HBP_n, useNA="ifany")


HBP <- ifelse(HBP_n==1|HBP_t==1,1, #This condition first
                ifelse(HBP_n==0|HBP_t==0,0,NA))

table(HBP, useNA="ifany")


```

			
			
### 3.13 Rheumatoid arthritis (RA)

Coding 'Do not know' as to other serios medical conditions also as 0 - difference from DW method

```{r}

RA_n <- num_recode(var1= "f.20002.0.0", #first var in range
                        var2 = "f.20002.3.33",#second var in range
                        code=1464) #Text to search in array and recode to NA)
table(RA_n, useNA="ifany")

#This Q for Other serious medical condition/disability diagnosed by doctor
noothercond <- text_recode0only(var1="f.2473.0.0",
                               var2="f.2473.3.0",
                               text="No|Do not know")

table(noothercond)

RA <- ifelse(RA_n==1, 1, #This condition first
                ifelse(RA_n==0|noothercond==0,0,NA))

table(RA, useNA="ifany")


```
		
### 3.14 Osteoarthritis 

disease code 1465 in field 20002

```{r}

osteoarthritis_n <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1465) #Text to search in array and recode to NA)
table(osteoarthritis_n, useNA="ifany")


osteoarthritis <- ifelse(osteoarthritis_n==1, 1, #This condition first
                ifelse(osteoarthritis_n==0|noothercond==0,0,NA))

table(osteoarthritis, useNA="ifany")


```

### 3.15 Gout 

disease code 1466 in field 20002

```{r}

gout_n <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1466) #Text to search in array and recode to NA)
table(gout_n, useNA="ifany")


gout <- ifelse(gout_n==1, 1, #This condition first
                ifelse(gout_n==0|noothercond==0,0,NA))

table(gout, useNA="ifany")


```


### 3.16 Dental problems

gen i16_dental = 0 if n_6149_0_0==-7
replace i16 = 1 if n_6149_0_0!=-7 & n_6149_0_0!=-3 & n_6149_0_0!=.

**Q** - could remove dentured and toothache potentially from this but 
keep for now

```{r}

dental.problems_t <- text_recode(var1= "f.6149.0.0", var2 = "f.6149.3.5",
                        text = "Bleeding gums|Dentures|Loose teeth|Mouth ulcers|Painful gums|Toothache", #Text to search in array and recode to 1
                        NAvals = "Prefer not to answer") 

table(dental.problems_t, useNA="ifany")

dental.problems <- ifelse(dental.problems_t==-99999, NA, dental.problems_t)
table(dental.problems, useNA="ifany")


```


### 3.17 Chest pain

Deviating from DW to code 'Do not know' as 0

```{r}
chest.pain_t <- text_recode(var1= "f.2335.0.0", var2 = "f.2335.3.0",
                        text = "Yes",
                        NAvals = "Prefer not to answer") 

table(chest.pain_t, useNA="ifany")

chest.pain <- ifelse(chest.pain_t==-99999, NA, chest.pain_t)
table(chest.pain, useNA="ifany")

```
  

### 3.18 Sciatica 
disease code 1476 in field 20002 

```{r}

sciatica_n <- num_recode(var1= "f.20002.0.0", #first var in range
                        var2 = "f.20002.3.33",#second var in range
                        code=1476) #Text to search in array and recode to NA)
table(sciatica_n, useNA="ifany")

sciatica <- ifelse(sciatica_n==1, 1, #This condition first
                ifelse(sciatica_n==0|noothercond==0,0,NA))

table(sciatica, useNA="ifany")
```

### 3.19 Diabetes
	
20002 codes 1220 and 1223 refer to 'diabetes' and 'type 2 diabetes' respectively 

Do not know coded as 0


```{r}
diabetes_t <- text_recode(var1= "f.2443.0.0", var2 = "f.2443.3.0",
                        text = "Yes", NAvals = "Prefer not to answer") 

table(diabetes_t, useNA="ifany")

#'1220 diabetes
diabetes_n1 <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1220)
table(diabetes_n1, useNA="ifany")

#'1223 T2 diabetes
diabetes_n2 <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1223) #Text to search in array and recode to NA)
table(diabetes_n2, useNA="ifany")

diabetes <- ifelse(diabetes_n1==1|diabetes_n2==1|diabetes_t==1,1, #This condition first
                ifelse(diabetes_n1==0|diabetes_n2==0|diabetes_t==0,0,NA))

table(diabetes, useNA="ifany")

```

### 3.20 Cancer


```{r}

cancer_t  <- text_recode(var1= "f.2453.0.0", var2 = "f.2453.3.0",
                        text = "Yes", NAvals = "Prefer not to answer") 

table(cancer_t, useNA="ifany")

#134 - number self reported cancers 
var1 = "f.134.0.0"
  var2 = "f.134.3.0"
  
  df.b <- df.a %>% select(var1:var2)
  
  df.b[is.na(df.b)] <- -99999

#Extracts the highest number across the array 
cancer_n_total <- apply(df.b,1, function(b){max(b)})

table(cancer_n_total, useNA="ifany")

cancer <- ifelse(cancer_n_total>=1|cancer_t==1,1, #This condition first
                ifelse(cancer_n_total==0|cancer_t==0,0,NA))

table(cancer, useNA="ifany")

```
### 3.21 Multiple cancers

```{r}

cancer.multi <- ifelse(cancer_n_total>1,1, #This condition first
                  ifelse(cancer_n_total<=1,0,NA))

table(cancer.multi, useNA="ifany")

```

### 3.22 Fractures

Again 'Do not know' coded as zero

```{r}

fractures_t <- text_recode(var1= "f.2463.0.0", var2 = "f.2463.3.0",
                        text = "Yes",
                        NAvals = "Prefer not to answer") 

table(fractures_t, useNA="ifany")

fractures <- ifelse(fractures_t==-99999, NA, fractures_t)
table(fractures, useNA="ifany")


```


### 3.23 Deep vein thrombosis (DVT)


```{r warning = F}

dvt_t <- text_recode(var1= "f.6152.0.0", var2 = "f.6152.3.4",
                        text = "DVT", NAvals = "Prefer not to answer") 

table(dvt_t, useNA="ifany")


dvt_n <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1094)
table(dvt_n, useNA="ifany")

dvt <- ifelse(dvt_n==1|dvt_t==1,1, #This condition first
                ifelse(dvt_n==0|dvt_t==0,0,NA))

table(dvt, useNA="ifany")

```


### 3.24 Emphysema/chronic bronchitis


```{r}
#emphysema is the same variable as CD 
emphysema_t <- text_recode(var1= "f.6152.0.0", var2 = "f.6152.3.4",
                        text = "Emphysema", NAvals = "Prefer not to answer") 

table(emphysema_t, useNA="ifany")


emphysema_n <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1113)
table(emphysema_n, useNA="ifany")

emphysema <- ifelse(emphysema_n==1|emphysema_t==1,1, #This condition first
                ifelse(emphysema_n==0|emphysema_t==0,0,NA))

table(emphysema, useNA="ifany")

```

### 3.25 Asthma

```{r}

asthma_t <- text_recode(var1= "f.6152.0.0", var2 = "f.6152.3.4",
                        text = "Asthma", NAvals = "Prefer not to answer") 

table(asthma_t, useNA="ifany")


asthma_n <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1111)
table(asthma_n, useNA="ifany")

asthma <- ifelse(asthma_n==1|asthma_t==1,1, #This condition first
                ifelse(asthma_n==0|asthma_t==0,0,NA))

table(asthma, useNA="ifany")

```


### 3.26 Allergies

Coded to include eczema and hayfever/rhinitis

```{r}

#grepl will pick up Hayfever, allergic rhinitis or eczema
allergies_t <- text_recode(var1= "f.6152.0.0", var2 = "f.6152.3.4",
                        text = "Hayfever", NAvals = "Prefer not to answer") 

table(allergies_t, useNA="ifany")

#1387 hayfever/rhinitis
allergies_n1 <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1387)
table(allergies_n1, useNA="ifany")

#1452 // eczema code
allergies_n2 <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1452)
table(allergies_n2, useNA="ifany")

allergies <- ifelse(allergies_n1==1|allergies_n2==1|allergies_t==1,1, #This condition first
                ifelse(allergies_n1==0|allergies_n2==0|allergies_t==0,0,NA))

table(allergies, useNA="ifany")


```

### 3.27 Hypothyroidism 

```{r}

hypothyrodism_n <- num_recode(var1= "f.20002.0.0", #first var in range
                        var2 = "f.20002.3.33",#second var in range
                        code=1226) #Text to search in array and recode to NA)
table(hypothyrodism_n, useNA="ifany")


hypothyrodism <- ifelse(hypothyrodism_n==1, 1, #This condition first
                ifelse(hypothyrodism_n==0|noothercond==0,0,NA))


table(hypothyrodism, useNA="ifany")

```

### 3.28 Recent depressed mood (depression)


Same Q re timings as in 3.5?
Also wonder if coding should be consistent with that? (0.5 here rather than 0.2?)

```{r}

 var1 = "f.2050.0.0"
  var2 = "f.2050.3.0"
  text0 = "Not at all"
  text0.5 = "Several days"
  text0.75 = "More than half the days"
  text1= "Nearly every day"
  NAvals = "Do not know|Prefer not to answer"
  
  df.b <- df.a %>% select(var1:var2)
  
  df.b[is.na(df.b)] <- -99999
  
depression_t <- apply(df.b,1, function(b){
    v <- ifelse(any(grepl(text1, b)), 1,
                ifelse(any(grepl(text0.75, b)), 0.75,
                       ifelse(any(grepl(text0.5, b)), 0.5,
                              ifelse(any(grepl(text0, b)), 0, -99999))))})

table(depression_t, useNA="ifany")

depression <- ifelse(depression_t==-99999, NA, depression_t)
table(depression, useNA="ifany")

```

### 3.29 Anxiousness 

Deviating from method to code 'Do not know' as 0

```{r}

anxiousness_t <- text_recode(var1= "f.1970.0.0", var2 = "f.1970.3.0",
                        text = "Yes",
                        NAvals = "Prefer not to answer") 

table(anxiousness_t, useNA="ifany")

anxiousness <- ifelse(anxiousness_t==-99999, NA, anxiousness_t)
table(anxiousness, useNA="ifany")


```

### 3.30 Severe anxiety 



```{r}

severe.anxiety_n <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=1111)
table(severe.anxiety_n, useNA="ifany")


severe.anxiety <- ifelse(severe.anxiety_n==1, 1, #This condition first
                ifelse(severe.anxiety_n==0|noothercond==0,0,NA))

table(severe.anxiety, useNA="ifany")


```


### 3.31 Mood - misery

```{r}

misery_t <- text_recode(var1= "f.1930.0.0", var2 = "f.1930.3.0",
                        text = "Yes",
                        NAvals = "Prefer not to answer") 

table(misery_t, useNA="ifany")

misery <- ifelse(misery_t==-99999, NA, misery_t)
table(misery, useNA="ifany")

```

### 3.32 Loneliness

```{r}

loneliness_t <- text_recode(var1= "f.2020.0.0", var2 = "f.2020.3.0",
                        text = "Yes",
                        NAvals = "Prefer not to answer") 

table(loneliness_t, useNA="ifany")

loneliness <- ifelse(loneliness_t==-99999, NA, loneliness_t)
table(loneliness, useNA="ifany")

```

### 3.33 Head and/or neck pain 

Assume head pain = headache

```{r}

head.neck.pain_t <- text_recode(var1= "f.6159.0.0", var2 = "f.6159.3.6",
                        text = "Neck or shoulder pain|Headache",
                        NAvals = "Prefer not to answer" ) 

table(head.neck.pain_t, useNA="ifany")

head.neck.pain <- ifelse(head.neck.pain_t==-99999, NA, head.neck.pain_t)
table(head.neck.pain, useNA="ifany")

```


### 3.34 to 3.39: Pain domains - (34) Back pain, (35) stomach pain, (36) Hip, (37) knee pain, (38) whole body, (39) facial pain

Same code so running in loop 

```{r}

t <- unique(df.a$f.6159.0.0)
t <- t[c(1,4,6:9)] #Pull out vars to be coded

pain.domainsL <- lapply(t, function(t){
  var_t <- text_recode(var1= "f.6159.0.0", var2 = "f.6159.3.6",
                        text = t,
                        NAvals = "Prefer not to answer" ) 
  
  var <- ifelse(var_t==-99999, NA, var_t)})

names(pain.domainsL) <- gsub(" ", ".", t)

lapply(pain.domainsL, function(x){table(x, useNA="ifany")})

list2env(pain.domainsL, .GlobalEnv)

```

### 3.40 Sleep
ACE touchscreen question "Do you have trouble falling asleep at night or do you wake up in the middle of the night?"


```{r}

var1 = "f.1200.0.0"
  var2 = "f.1200.3.0"
  text0 = "Never/rarely"
  text0.5 = "Sometimes"
  text1= "Usually"
  NAvals = "Prefer not to answer"
  
  df.b <- df.a %>% select(var1:var2)
  
  df.b[is.na(df.b)] <- -99999
  
sleep_t <- apply(df.b,1, function(b){
    v <- ifelse(any(grepl(text1, b)), 1,
                ifelse(any(grepl(text0.5, b)), 0.5,
                              ifelse(any(grepl(text0, b)), 0, -99999)))})

table(sleep_t, useNA="ifany")

sleep <- ifelse(sleep_t==-99999, NA, sleep_t)
table(sleep, useNA="ifany")


```


### 3.41 High cholesterol

*note* - need to combine separate vars for male and female participants - based on medication use

	* female vars n_6153_0; male n_6177_0

gen i41_cholesterol =  0 if n_6153_0_0!=1 & n_6177_0_0!=1
replace i41 = 1 if n_6153_0_0==1
replace i41 = 1 if n_6177_0_0==1

	* no cholesterol therapeutic uses recorded in other array vars 

replace i41= . if n_6153_0_0==-3 |  n_6177_0_0==-3
replace i41= . if n_6153_0_0==-1 |  n_6177_0_0==-1

```{r}

#female var
high.cholesterol_t1 <- text_recode(var1= "f.6153.0.0", var2 = "f.6153.3.3",
                        text = "Cholesterol",
                        NAvals = "Prefer not to answer") 

table(high.cholesterol_t1, useNA="ifany")

#male var
high.cholesterol_t2 <- text_recode(var1= "f.6177.0.0", var2 = "f.6177.3.2",
                        text = "Cholesterol",
                        NAvals = "Prefer not to answer") 

table(high.cholesterol_t2, useNA="ifany")

high.cholesterol <- ifelse(high.cholesterol_t1==1|high.cholesterol_t2==1, 1,
                           ifelse(high.cholesterol_t1==0|high.cholesterol_t2==0,0,NA))
                     
table(high.cholesterol, useNA="ifany")


```

### 3.42 - 3.49 Other conditions 

All same coding (and as same as other above but easier to keep track this way):
 
* 42 Pneumonia - code 1398 in 20002
* 43 Gastric reflux - code 1138 in 20002
* 44 Hiatus hernia - code 1474 in 20002
* 45 diverticulitis - code 1458 in 20002
* 46 Gall stones - code 1162 in 20002
* 47 Psoriasis - code 1453 in 20002
* 48 Osteoporosis - code 1309 in 20002
* 49 Migraine - code 1265 in 20002 


```{r}

codes <- c(1398, 1138, 1474, 1458, 1162, 1453, 1309, 1265)

other.conds.L <- lapply(codes, function(i){

      var_n <- num_recode(var1= "f.20002.0.0", var2 = "f.20002.3.33",
                        code=i)
  
      var <- ifelse(var_n==1, 1, #This condition first
                ifelse(var_n==0|noothercond==0,0,NA))})

names(other.conds.L) <- c("pneumonia", "gastric.reflux", "hiatus.hernia",
                          "diverticulitis", "gall.stones", "psoriasis",
                          "osteoporosis", "migrane")

lapply(other.conds.L, function(x){table(x, useNA="ifany")})

list2env(other.conds.L, .GlobalEnv)

```
## Generate FI

### Combine domains 

#here finsh
```{r}

FI.domains <- data.frame(id=df.a$f.eid, D1.glaucoma=glaucoma,
                         D2.cataracts=cataracts, D3.hearing=hearing,
                         D4.infirmity=infirmity, D5.self.rated.health=SRH,
                         D6.falls=falls, D7.wheeze=wheeze,
                         D8.fatigue=fatigue, D9.MI=MI, D10.angina=angina,
                         D11.stroke=stroke, D12.HBP=HBP, D13.RA=RA,
                         D14.osteoarthritis=osteoarthritis, D15.gout=gout,
                         D16.dental.probelms=dental.problems,
                         D17.chest.pain=chest.pain, D18.sciatica=sciatica,
                         D19.diabetes=diabetes, D20.cancer=cancer,
                         D21.cancer.multi=cancer.multi, 
                         D22.fractures=fractures, D23.DVT=dvt,
                         D24.emphysema=emphysema, D25.asthma=asthma,
                         D26.hypothyroidism=hypothyroidsi
                        )

```

```{r}

* Deriving FI - sum up items, divide by 49, and multiply by 100 (to display as percentages) or 10 (for results displayed by 10% increases)

mark all_items
markout all_items i*

egen total_points = rowtotal(i1_-i49_) if all_items

gen fi_49item = (total_points/49)

lab var fi "Frailty Index with 49 items - complete case set"


* Deriving missing data count for items in those with incomplete data:

egen missing_item_count = rowmiss(i1_-i49_) 

count if  missing_item_count>=10
		
		*   N=2,292 with data missing on 10 or more items - to be dropped

drop if missing>=10 // Leaves N= 500339 


	* Scaling FI to be on 1 and 10 percent increases (0.01 and 0.1 unit increments on FI scale)
gen fi_percent = fi_49item * 100

gen fi_10percent = fi_49item * 10


* categorised FI, 0.1 increments but combining all individuals with FI over 40:

egen ficat_10percent = cut(fi_percent), at(0,10,20,30,40,60)


/* Extra script for deriving a 'denominator-varying' FI - 
 * A commonly used approach to provide an FI on full samples where missing item data are present. 
 * This calculation divides summed deficits for items with available data by denominators that vary depending on the available item number

egen total_points_dvFI = rowtotal(i1_-i49_) 

gen fi_dv = total_points_dvFI/(49-missing_item_count)

lab var fi_dv "Frailty Index for whole cohort using total-varying denominators"

*/


******* Preparing ethnicity into a variable with limited number of categories (as per groupings on UKB Data Showcase)

gen ethnicity = . if n_21000_0_0 < 0

	* White category:
replace ethnicity = 1 if n_21000_0_0 == 1  
replace ethnicity = 1 if n_21000_0_0 >= 100 &  n_21000_0_0 < 2000

	* Asian category:
replace ethnic = 2 if n_21000_0_0 >2004 & n_21000_0_0<4000
replace ethnic = 2 if n_21000_0_0 == 3

	* Black category:
replace ethnic = 3 if n_21000_0_0>4000 & n_21000_0_0!=.
replace ethnic = 3 if n_21000_0_0 == 4

	* Mixed:
replace ethnic = 4 if n_21000_0_0>2000 & n_21000_0_0<3000
replace ethnic = 4 if n_21000_0_0==2

	* Chinese
replace ethnic = 5 if n_21000_0_0==5

	* Other
replace ethnic = 6 if n_21000_0_0==6

label define ethnic_groups 1 "White" 2 "Asian or Asian British" 3 "Black or Black British" 4 "Mixed" 5 "Chinese" 6 "Other"
label values ethnicity ethnic_groups



*** Death indicator:

gen died = 1 if n_40007_0_0!=.
replace died = 0 if n_40007_0_0==.

	* Recoding indicator for those that died after overall censorship date....
		/*Current UKB censoring ends Nov 2015 - some deaths are recorded from particular death stat sources after that, 
		but coverage isn't complete for all participants. Thus, any deaths at that date need replacing to 0 as if censored
		NB - more info on UKB censoring dates here: http://biobank.ctsu.ox.ac.uk/showcase/exinfo.cgi?src=Data_providers_and_dates */

di date("20151130","YMD") 
replace died = 0 if ts_40000_0_0>20422 & ts_40000_0_0!=.


*** Coding a time variable based on attained age = age at death or censorship - age at baseline


* Coding age at censorship for survivors:

gen censor_date = date("30/11/2015", "DMY") // Current overall censorship date - 30th Nov, 2015
replace censor_date = ts_40000_0_0 if died==1
format censor_date %d 

gen time_to_censor = (censor_date - ts_53_0_0) 

/* NB: returns three negative values - these individuals have death dates prior to their attendance dates;
	 ID numbers: 3077398 5373093 3448373
	One participant (5373093) has an alternate instance for death date (presumably where a second death certificate was issued, perhaps after post-mortem, as suggested in UKB death stats doc)
	HOWEVER - email from UKB Data Access team confirmed that all three should be dropped from analyses
*/


* Dropping the individuals with death dates before attendance dates:
drop if time_to_censor<0
	
	

* Variable that indicates the age at either censorship or death:

gen time_to_censor_years = (time_to_censor)/365.25
gen age_censor = n_21003_0_0 + time_to_censor_years


* coding a date of birth variable from month and year vars. All individuals assigned 15 as day of month born. 

gen day_born = 15
gen birth_date = mdy(n_52_0_0, day_born, n_34_0_0) 
format birth_date %d 


/* Setting stata into survival modelling mode,

- Time variable - Date of death or censoring
- Origin point - birth date (NB: approx b-date coded below from only month and year of birth vars available)
- Entry point: date at baseline assessment 
- Exit: death or censoring date, 30th Nov 2015

*/

stset censor_date, id(n_eid) failure(died==1) origin(birth_date) entry(ts_53_0_0) exit(time mdy(11,30,2015)) scale(365.25)



* Survival analyses on complete-case data:
	* Model 1 (adjusted for age only):
stcox fi_percent
stcox fi_10percent

	* Model 2:
stcox fi_percent n_31 
stcox fi_10percent n_31 

	* Model 3:
xi: stcox fi_percent n_31 i.ethnic
xi: stcox fi_10percent n_31 i.ethnic

* Checking PH assumption by log-log plot
stphplot, by(ficat_10percent)

* Checking PH by Schoenfeld: 
xi: stcox fi_10percent n_31 i.ethnic
estat phtest, detail
estat phtest, plot(fi_10percent) lineopts(lc(red))


* Hazards plot by FI category:
sts graph, hazard by(ficat_10percent) 

* Unadjusted survival curve by FI category
sts graph, by(ficat_10percent)  noorigin yla(.5(.1)1)

	
 * N-A cumulative hazards plots:
sts graph, cumhaz by(ficat_10percent)

 
* Generating the cumulative hazards function (nelson-aalen estimator) for including in MI input varlist:
sts gen cumul_haz = na




********************* Imputation ****************************

set seed 1066

* Test var set:

keep n_eid i* ///
cumul_haz died birth_date ts_53_0_0 censor_date n_21003_0_0 ethnicity n_31_0_0 _st _d _origin _t _t0

* Setting program for MI:
mi set mlong 


* Registering variables to be imputed:

	* Any with missing data:
mi register imputed i1_glaucoma i2_cataracts i3_hearing i4_infirmity i5_self_health ///
i6_falls i7_wheeze i8_fatigue i9_chd i10_angina i11_stroke i12_hypertension  ///
 i16_dental i17_chest_pain  i19_diabetes i20_cancer i21_multi_cancer i22_fractures /// 
i23_dvt i24_emphysema_cbronc i25_asthma i26_allergies  i28_depression i29_anxiousness ///
i31_misery i32_loneliness i33_headneck_pain i34_back_pain i35_stomach_pain ///
i36_hip_pain i37_knee_pain i38_wholebody_pain i39_facial_pain i40_sleep i41_cholesterol ethnicity


	* Vars/items with complete data
mi register regular n_eid cumul_haz died birth_date ts_53_0_0 censor_date n_21003_0_0 n_31_0_0 _st _d _origin _t _t0 ///
i13_rh_arth i14_osteoarth i15_gout i18_sciatica i27_hypothyroidism i30_severe_anxiety i42_pneumonia i43_gastric_reflux ///
i44_hernia i45_diverticulitis i46_gall_stones i47_psoriasis i48_osteoporosis i49_migraine



/* Multivariate imputation settings for vars of multiple types:

* List of binary vars to impute with logit:
i1_ i2_ i3_ i4_  i7 i9 i10 i11 i12 i13 i14 i15 i16 i17 i18 ///
i19 i20 i21 i22 i23 i24 i25 i26 i27 i29 i30 i31 i32 i33 i34 i35 ///
i36 i37 i38 i39 i41 i42 i34 i44 i45  i46 i47 i48 i49


* List of categorical vars (with uneven spacings) to impute with PMM:
i5 i6 i8 i28 i40 ethnicity


*/



***** MI command (NB: recommend running in parallel via Stata-MP with high-performance computing):

mi impute chained (logit, augment) i1_ i2_ i3_ i4_  i7 i9 i10 i11 i12  i16 i17 ///
i19 i20 i21 i22 i23 i24 i25 i26 i29 i31 i32 i33 i34 i35 i36 i37 i38 i39  i41 ///
(pmm, knn(5)) ethnicity i5 i6 i8 i28 i40 = _t0 _t _d cumul_haz n_31_0_0 i13 i14 i15 i18 ///
 i27 i30 i42 i43 i44 i45 i46 i47 i48 i49, add(5) noisily 


********* Analysis in full MI dataset with containing 5 imputations:


* Coding of FI within imputed data..... ran these commands and then saved a new set to avoid waiting for long variable generation procedures every time the set is loaded:


sort n_e _mi_m


* NB: data should be set in flong so that each imputation set has complete values for all, not just for those with missing values as in mlong style
mi convert flong

* creating FI in imputed data:

mi passive: egen miss_items = rowmiss(i*)
mi passive: egen total_points = rowtotal(i*) if miss_items==0 

	* it is necessary generate fi for all individuals with passive commands, rather than include FI var in imputation data and then use standard coding (the FI vals in imputed data get recoded to . in subsequent estimate commands)

mi passive: gen fi_49item = (total_points / 49)
mi passive: gen fi_percent = fi_49item*100
mi passive: gen fi_10percent = fi_49item*10
mi passive: egen ficat = cut(fi_percent), at(0,10,20,30,40,60)


* Age categories for examining FI distribution:
mi passive: egen agecats = cut(n_21003_0_0), at(35,50,60,65,75) icodes label


* Frac poly of FI on age:

fp <n_21003_0_0>: regress fi_percent <n_21003_0_0> if _mi_m==2

* FP Plot:
twoway scatter fi_49item _t0 if _mi_m==1, mfc(none) msiz(vsmall) mlc(gs8%20) || fpfit fi_49item _t0 if _mi_m==1, lcolor(gs0) lwidth(medium) ///
legend(label(1 "Observations") label(2 "Best fitting trend")) xlabel(40(10)75)  ///
xtitle("Age at baseline assessment") ytitle("Frailty Index values")


/* Testing frac poly models on other imputation sets and complete-case:
fp <n_21003_0_0>, replace: regress fi_percent <n_21003_0_0> if _mi_m==2
fp <n_21003_0_0>, replace: regress fi_percent <n_21003_0_0> if _mi_m==3
fp <n_21003_0_0>, replace: regress fi_percent <n_21003_0_0> if _mi_m==4
fp <n_21003_0_0>, replace: regress fi_percent <n_21003_0_0> if _mi_m==5

fp <n_21003_0_0>, replace: regress fi_percent <n_21003_0_0> if _mi_m==0

*/


** Survival models in MI:
	
	
****** Full set
mi stset censor_date, id(n_eid) failure(died==1) origin(birth_date) entry(ts_53_0_0) exit(time mdy(11,30,2015)) scale(365.25)


	* Model 1:
mi estimate: stcox fi_10percent
mi estimate, hr // need to add this post-estimate command to convert coeffs to HRs 

	* Model 2:
mi estimate:  stcox fi_10percent n_31 
mi estimate, hr 

	* Model 3:
mi estimate:  stcox fi_10percent n_31 i.ethnic
mi estimate, hr 



******* By sex:
	
xi: stcox i.n_31*fi_10percent if _mi_m==0 // interaction tested in complete-case data - P<0.001


 * Model 1: 
mi estimate: stcox fi_10percent if n_31==0
mi estimate, hr 

mi estimate: stcox fi_10percent if n_31==1
mi estimate, hr 

 * Model 3 (age and ethnicity)
mi estimate: stcox fi_10percent i.ethnic if n_31==0
mi estimate, hr 

mi estimate: stcox fi_10percent i.ethnic if n_31==1
mi estimate, hr 

	

* FI vs survival by baseline age groups:


foreach n of numlist 0(1)3 {

mi estimate: stcox fi_10percent if agecats==`n' 
mi estimate, hr 

mi estimate: stcox fi_10percent n_31 if agecats==`n' 
mi estimate, hr 

mi estimate: stcox fi_10percent n_31 i.ethnic if agecats==`n' 
mi estimate, hr 

}

	* repeat in complete-case:
foreach n of numlist 0(1)3 {

stcox fi_10percent if agecats==`n' & _mi_m==0

 stcox fi_10percent n_31 if agecats==`n' & _mi_m==0

xi: stcox fi_10percent n_31 i.ethnic if agecats==`n' & _mi_m==0


}


**** Averaging K-M curves from separate imputed sets (no official command for this):

	* Plotting based on FI averaged over five imputaitons:
	
mi convert wide, clear

gen fi_average = (_1_fi_49item+_2_fi_49item+_3_fi_49item+_4_fi_49item+_5_fi_49item)/5
gen fi_av_percent = fi_ave * 100
egen fi_av_cat = cut( fi_av_percent), at(0,10,20,30,40,60)
	
sts graph if _mi_m==0, by(fi_av_cat)  noorigin yla(.5(.1)1) title("") ///
ylab(0.55(0.1)1, format(%12.1fc)) ///
	plot1opts(lpattern(solid) lcolor(black)) ///
	plot2opts(lpattern(longdash_dot) lcolor(black)) ///
	plot3opts(lpattern(shortdash) lcolor(black)) ///
	plot4opts(lpattern(dash_dot) lcolor(black)) ///
	plot5opts(lpattern(dash) lcolor(black)) ///
legend( subtitle("FI categories") label(1 "< 0.1") label(2 "0.1 to < 0.3") label(3 "0.2 to < 0.3") label(4 "0.3 to < 0.4") label(5 "≥ 0.4")) ///
xtitle("Attained age") ytitle("Survival probability")


* Histogram based on averaged FI:

hist fi_average, color(gs13) lcolor(black) percent

lab var fi_average "Frailty Index score - proportion of deficits accrued"
twoway (histogram fi_average if n_31==0, bin(20)  percent color(gs13)) ///
       (histogram fi_average if n_31==1, bin(20) percent   ///
	   fcolor(none) lcolor(black)), legend(order(1 "Female" 2 "Male" ))
	   




# MICE 

https://www.r-bloggers.com/2016/06/handling-missing-data-with-mice-package-a-simple-approach/
