---
title: "Recreating UKBB FI in R"
author: "Ruth CE Bowyer"
date: "2023-08-30"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
```

## 0. About

Converting the Stata script provided by Williams et al., to generate an FI in UKBB

Differences are highlighted below but in brief:


This script was run mostly via CREATE

```{r}

library(dplyr)

```

## 1. Data

Variables to select as per William's et al.,:

Variables required from UKB data (these represent data field numbers, along with the Stata prefix for each variable 
	('s_' or 'n_' for string or numeric vars, respectively):
n_eid  		-- participant identifier x
n_31_0* 	-- Sex x
n_34_* 		-- Year of birth x
n_52_* 		-- Month of birth x
n_21003_0*  -- Age when attended assessment centre x
n_21000* 	-- Ethnicity x
n_1200_0* 	-- Sleeplessness / insomnia x
n_6148_0* 	-- Eye problems / disorders x
n_2227_0*  	-- Other eye problems x
n_2020_0* 	-- Loneliness / isolation x
n_2247_0* 	-- Hearing difficulties x
n_2188_0* 	-- Long-standing illness, disability or infirmity x
n_2178_0* 	-- Overall self-rated health x
n_2296_0* 	-- Falls in last year x
n_2316_0* 	-- Chest wheeze or whistling x
n_2080_0* 	-- Tiredness / Lethargy x
n_6150_0* 	-- Vascular problems	x
n_20002_0* 	-- Non-cancer illness codes (all array variables are necessary to keep) x
n_6149_0* 	-- Mouth / dental problems x 
n_2335_0* 	-- Chest pain x
n_2443_0* 	-- Diabetes x
n_2453_0* 	-- Cancer diagnosis x
n_134_0* 	-- Number of self-reported cancers x
n_2463_0* 	-- fractured/broken bones x
n_6152_0*	-- Multiple medical conditions reported - e.g. DVT, asthma, eczema x
n_3786_0*	-- Age of asthma diagnosis x
n_2473_0* 	-- Other serious medical conditions reported x
n_2050_0*  	-- Depressed moods x
n_1970_0* 	-- Nervous feelings x
n_1930_0* 	-- Miserableness x
n_6159_0* 	-- Pain in last month x
n_136_0* 	-- Number of operations x

n_6177_0* 	-- Medications used for cholesterol, blood pressure or diabetes (male)
n_6153_0* 	-- Medications used for cholesterol, blood pressure or diabetes (female)
n_137_0* 	-- Number of medications used

n_40007_* 	-- Age at death for deceased participants

```{r eval=FALSE, include=T}
# This chunk ran for set up and then just loaded as below for subsequent iterations
#df.a derived as follows: 

load("~/max2ruth2/bd671644_recoded.Robj")

df.a <- bd %>% select("f.eid", "f.31.0.0", "f.34.0.0", "f.52.0.0",
                      "f.21003.0.0":"f.21003.3.0", "f.21000.0.0":"f.21000.3.0",
                      "f.1200.0.0":"f.1200.3.0", "f.6148.0.0":"f.6148.3.4", 
                      "f.2227.0.0":"f.2227.3.0", "f.2020.0.0":"f.2020.3.0",
                      "f.2188.0.0":"f.2188.3.0", "f.2178.0.0":"f.2178.3.0",
                      "f.2296.0.0":"f.2296.3.0", "f.2316.0.0":"f.2316.3.0",
                      "f.2080.0.0":"f.2080.3.0", "f.6150.0.0":"f.6150.2.3",
                      "f.20002.0.0":"f.20002.3.33", "f.6149.0.0":"f.6149.3.5",
                      "f.2335.0.0":"f.2335.3.0", "f.2443.0.0":"f.2443.3.0",
                      "f.2453.0.0":"f.2453.3.0", "f.134.0.0":"f.134.3.0",
                      "f.2463.0.0":"f.2463.3.0", "f.6145.0.0":"f.6145.3.2",
                      "f.3786.0.0":"f.3786.3.0", "f.2473.0.0":"f.2473.3.0", "f.2050.0.0":"f.2050.3.0",
                      "f.1970.0.0":"f.1970.3.0", "f.1930.0.0":"f.1930.3.0", "f.6159.0.0":"f.6159.3.6",
                      "f.136.0.0":"f.136.3.0", "f.6177.0.0":"f.6177.3.2", "f.6153.0.0":"f.6153.3.3",
                      "f.137.0.0":"f.137.3.0", "f.40007.0.0":"f.40007.1.0")

write.csv(df.a,"/users/k1511199/UKBB_multimorbidity/UKBB.df.a.csv", row.names=F)

rename.df <- names(df.a)

```

Many of the array vals seem to be completely missing NA 

```{r}

df.a <- read.csv("/users/k1511199/UKBB_multimorbidity/UKBB.df.a.csv")
## NA those who withdrew 
x<-read.csv("~/max2ruth/w90865_2023-04-25.csv",h=F)

df.a <- df.a[-which(df.a$f.eid%in%x$V1),]

## set row names as study id for sense checking

row.names(df.a) <- paste0("id",df.a$f.eid)
```


## 2. Write functions 


```{r recode function}

#Function to search within a range of cols for a text variable and recode as 1 - returns a vector of the var 
text_recode <- function(var1, #first var in range
                        var2 = NULL, #second var in range
                        text, #Text to search in array and recode to NA
                        NAvals, ...){ #Text to recode, sep multiple with |
  var1 = var1
  var2 = var2
  text = text
  df.b <- df.a %>% select(var1:var2)
  
  #Codes any mention of morb as 1, and if no 1s and all responses arent NA, returns 0
  var <- apply(df.b,1, function(b){
    v <- ifelse(any(grepl(text, b)), 1, 
           ifelse(all(grepl(NAvals, b)|is.na(b)), NA, 0))})
  return(var)}


num_recode <- function(var1, #first var in range
                        var2 = NULL, #second var in range
                        code){ #Code for morb 
  var1 = var1
  var2 = var2
  code = code
  df.b <- df.a %>% select(var1:var2)
  
  #NAs really preventing the ifelse from running for reasons unable to work out, so adding here a workaround (unclear to me why ifelse(all(is.na(b))) is not working!)

  df.b[is.na(df.b)] <- -11111111
  var <- apply(as.data.frame(df.b),1,function(b){
    b <- as.numeric(b)
    ifelse(all(b==-11111111), NA, 
               ifelse(any(b==code),1,0))
  })
  
  return(var)}


```

## 3. Clean datasets 


### 3a. Glaucoma

- Within f.6148* - text var
- is code 1277 in 20002 
- Recode do not know and pref not to answer to NA 

Questions from og methodology: 

Var 2227 is other eye problems (answers no, prefer to not answer, yes) -- seems to be using this first but i am not sure why? As other eye problems could be diff to? (Not included for now)

I will recode pref not to answer and don't know to NA as in the og -- but im not sure this is correct given the MICE? (and some don't knows could be recoded to 0)

'	* recoding 6148 'do not know' respondents to 0 if they were interviewed and still had no record of glaucoma:
replace i1_glaucoma = 0 if i1_!=1 & n_20002_0_0!=. & n_6148_0_0==-1 '

I'm not sure what variable is meant by this or the variable i1 - which i think is to do with interviewing but I can't see a corresponding var in the html, or which variable this relates to 

99999 in 20002 -- unclassifable so I don't think coding as NA is appropiate but discuss
https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=6

Where all responses in 20002 is NA - the fnc codes as NA(but could be 0 if the NAs are responses vs not being asked)


```{r}

glauc_t <- text_recode(var1= "f.6148.0.0", #first var in range
                        var2 = "f.6148.3.4",#second var in range
                        text = "Glaucoma", #Text to search in array and recode to NA
                        NAvals = "Prefer not to answer|Do not know") #Text to recode, sep multiple with 

table(glauc_t, useNA="ifany")

glauc_n <- num_recode(var1= "f.20002.0.0", #first var in range
                        var2 = "f.20002.3.33",#second var in range
                        code=1277) #Text to search in array and recode to NA)
table(glauc_n, useNA="ifany")


glauc <- ifelse(glauc_n==1|glauc_t==1,1,
                ifelse(glauc_n==0|glauc_t==0,0,NA))

table(glauc, useNA="ifany")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Generate FI

# MICE 

https://www.r-bloggers.com/2016/06/handling-missing-data-with-mice-package-a-simple-approach/
